### 2020-04-20 final_report/methodology.tex
minor edits in the metrics section, change f tilde to f hat

Labeling the expensive Monte Carlo simulation $f(x)$, a surrogate is a mapping
$\tilde{f}(x)$ that yields similar images as $f(x)$. In other words, $f(x)$ and
$\tilde{f}(x)$ minimise a selected similarity metric. Furthermore, in order to
be considered \textit{viable}, surrogates are required to achieve expected evaluation time
that does not exceed the expected evaluation time of $f(x)$.

In the supervised learning setting, we first gather a sufficiently large
training set of samples $\mathcal{T}=\left\{\left( x^{(i)},f\left(x^{(i)}\right) \right)\right\}_{i=1}^N$
to describe the behaviour of $f(x)$ across its domain.
Depending on specific model class and appropriate choice of its
hyperparameters, surrogate models $\tilde{f}(x)$ are trained to minimise
empirical risk with respect to $\mathcal{T}$ and a model-specific
loss function $\mathcal{L}$, where empirical risk is defined as

\begin{align}
	R_{\text{emp.}}(\tilde{f}\mid\mathcal{T},\mathcal{L})
	=\frac{1}{N}\sum_{i=1}^N
	\mathcal{L}\left(\tilde{f}(x^{(i)}),f(x^{(i)})\right).
\end{align}

The unsupervised setting can be viewed as an extension of this procedure.
Rather than fixing the training set $\mathcal{T}$ for the entire duration of
training, the points of evaluation $\{x^{(i)}\}_{i=1}^N$ that determine the set
are first initialised randomly, and continuously extended throughout training. This
permits the learning algorithm to motivate the choice of new points following the
evaluation of surrogates trained thus far by appropriately biasing the
proposal distribution, in order to better focus on problematic regions within
the domain.

>>>>

Labeling the expensive Monte Carlo simulation $f(x)$, a surrogate is a mapping
$\hat{f}(x)$ that yields similar images as $f(x)$. In other words, $f(x)$ and
$\hat{f}(x)$ minimise a selected similarity metric. Furthermore, in order to
be considered \textit{viable}, surrogates are required to achieve expected evaluation time
that does not exceed the expected evaluation time of $f(x)$.

In the supervised learning setting, we first gather a sufficiently large
training set of samples $\mathcal{T}=\left\{\left( x^{(i)},f\left(x^{(i)}\right) \right)\right\}_{i=1}^N$
to describe the behaviour of $f(x)$ across its domain.
Depending on specific model class and appropriate choice of its
hyperparameters, surrogate models $\hat{f}(x)$ are trained to minimise
empirical risk with respect to $\mathcal{T}$ and a model-specific
loss function $\mathcal{L}$, where empirical risk is defined as

\begin{align}
	R_{\text{emp.}}(\hat{f}\mid\mathcal{T},\mathcal{L})
	=\frac{1}{N}\sum_{i=1}^N
	\mathcal{L}\left(\hat{f}(x^{(i)}),f(x^{(i)})\right).
\end{align}

The unsupervised setting can be viewed as an extension of this method.
Rather than fixing the training set $\mathcal{T}$ for the entire duration of
training, multiple sets $\{\mathcal{T}_k\}_{k=0}^K$ are used, such that
$\mathcal{T}_{k-1}\subset\mathcal{T}_k$ for all $k>1$. The first set
$\mathcal{T}_0$ is initialised randomly to provide a \textit{burn-in}, and is
repeatedly extended in epochs, whereby each epoch trains a new surrogate on
$\mathcal{T}_k$ using the supervised learning procedure, evaluates its
performance, and forms a new set $\mathcal{T}_{k+1}$ by adding more samples to
$\mathcal{T}_k$. This permits the learning algorithm to condition the selection
of new samples by the results of evaluation in order to focus on improvement of
surrogate performance in complex regions within the domain.

\subsection{Metrics}
\label{sec:metrics}

Aiming to provide objective comparison of a diverse set of surrogate model
classes, we define a multitude of metrics to be tracked during experiments.
Following the motivation of this work, two desirable properties of surrogates
arise: (i) their capability to approximate the expensive
model well and (ii) their time of evaluation. An ideal surrogate would maximise
the former while minimising the latter.

To prevent undesirable bias in results due to training set selection, metrics
related to both properties are collected using $k$-fold cross-validation with a
standard choice of $k=5$. Herein, a sample set is subdivided into 5 disjoint
folds which are repeatedly interpreted as training and testing sets, maintaining
a constant ratio of samples between the two. In each such interpretation experiments are
repeated, and the overall value of each metric of interest is reported as the mean
across all folds.

\Cref{tbl:metrics} provides exhaustive list and description of
regression performance and evaluation time metrics recorded in the experiments.

\begin{table}[h]
	\centering
	\begin{tabular}{lll}
	\toprule
	Regression performance metrics	& Mathematical formulation / description & Ideal value \\
	\midrule
	Mean absolute error	& $\sum_{i=1}^N |y^{(i)}-\hat{y}^{(i)}|/N$ & 0 [TBR] \\
	$R^2$ ratio (coefficient of determination)	& $1-\sum_{i=1}^N \left(y^{(i)}-\hat{y}^{(i)} \right)^2 /
	\sum_{i=1}^N \left( y^{(i)}-\overline{y} \right)^2 $ & 1 [rel.] \\
	Adjusted $R^2$ ratio	& $1-(1-R^2)(N-1)/(N-P-1)$	& 1 [rel.] \\
	Standard error of regression	& $\text{StdDev}_{i=1}^N\left( |y^{(i)} -
	\hat{y}^{(i)}| \right) $	 & 0 [TBR] \\
	\midrule
	Evaluation time metrics	& {} & {} \\
	\midrule
	Mean sample training time	& $(\text{wall training time of $\hat{f}(x)$})/N_0$ 	& 0 [ms] \\
	Mean sample prediction time	& $(\text{wall evaluation time of $\hat{f}(x)$})/N$	& 0 [ms] \\
	\bottomrule
	\end{tabular}
	\caption{Metrics recorded in supervised learning experiments. In
	formulations, we work with training set of size $N_0$ and testing set of
size $N$, TBR values $y^{(i)}=f(x^{(i)})$ and $\hat{y}^{(i)}=\hat{f}(x^{(i)})$
denote images of the $i$th sample in the expensive model and the surrogate
respectively. Furthermore, the mean $\overline{y}=\sum_{i=1}^N y^{(i)}/N$ and $P$ is the
number of input features.}
	\label{tbl:metrics}
\end{table}
